---
title: "COTS Population Model: explore data"
author: "Sam Matthew, Camille Mellin, Damien Fordham, Kevin Shoemaker, ..."
date: "April 7, 2016"
output: html_document
---

# Prepare the R workspace (code running behind the scenes- see the .Rmd file) 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)    # set up knitr

#######################
#   CLEAR THE WORKSPACE 
#######################

rm(list=ls())

#######################
#   SET USER 
#######################

USER = "KEVIN"

#########################
# SET PROJECT DIRECTORIES (this should be the only place where local directories should be referenced)
#########################

if(USER=="KEVIN") BASE_DIRECTORY <- "C:\\Users\\Kevin\\Dropbox\\CoTS_Model"             # NOTE: this should link to the Dropbox folder with shared project resources	                                                                        
if(USER=="KEVIN") CODE_DIRECTORY <- "C:\\Users\\Kevin\\GIT\\COTS_popmodel"              # NOTE: code directory should be your local copy of the GitHub repository

if(USER=="SAM") BASE_DIRECTORY ##<- ****FILL IN*****
if(USER=="SAM") CODE_DIRECTORY ##<- ****FILL IN*****

SPATIALDATA_DIRECTORY <- paste(BASE_DIRECTORY,"\\Spatial Layers",sep="")                          # directory for storing relevant spatial data (ASC, SHP files)
if(is.na(file.info(SPATIALDATA_DIRECTORY)[1,"isdir"])) dir.create(SPATIALDATA_DIRECTORY)

DATA_DIRECTORY <- paste(BASE_DIRECTORY,"\\Data",sep="")                                 # directory for storing data (CSV files)
if(is.na(file.info(DATA_DIRECTORY)[1,"isdir"])) dir.create(DATA_DIRECTORY)

ENVDATA_DIRECTORY <- paste(DATA_DIRECTORY,"\\Environmental",sep="")                                 # directory for storing data (CSV files)
if(is.na(file.info(ENVDATA_DIRECTORY)[1,"isdir"])) dir.create(ENVDATA_DIRECTORY)

FIGURES_DIRECTORY <- paste(BASE_DIRECTORY,"\\Figures\\RawFigures",sep="")               # directory for storing raw figures generated from R
if(is.na(file.info(FIGURES_DIRECTORY)[1,"isdir"])) dir.create(FIGURES_DIRECTORY)

RESULTS_DIRECTORY <- paste(BASE_DIRECTORY,"\\results",sep="")                           # directory for storing relevant results
if(is.na(file.info(RESULTS_DIRECTORY)[1,"isdir"])) dir.create(RESULTS_DIRECTORY)

RDATA_DIRECTORY <- paste(BASE_DIRECTORY,"\\R_Workspaces",sep="")                        # directory for storing .RData files (R workspaces and data objects)
if(is.na(file.info(RDATA_DIRECTORY)[1,"isdir"])) dir.create(RDATA_DIRECTORY)

####################
#  LOAD FUNCTIONS AND SCRIPTS FROM SOURCE CODE (your local GitHub repository)
#################### 

setwd(CODE_DIRECTORY)
source("COTSModel_Utilityfunctions.R")   # load utility functions, e.g., for loading packages etc. 
#source("COTSModel_COTSfunctions.R")      # load functions for implementing COTS demography and dispersal
#source("COTSModel_Coralfunctions.R")     # load functions for implemeting coral growth/recovery and dispersal
source("COTSModel_GISfunctions.R")     # load functions for implemeting coral growth/recovery and dispersal
source("COTSModel_Initialization.R")     # initialize the workspace

```

```{r echo=FALSE}
cat(sprintf("The current user is %s",USER))
```


# COTS Population Model: objectives

This population model is constructed in R, using the [RStudio](www.rstudio.com) development environment. 
Yadda yadda **yadda** *yadda* ~~yadda~~ [yadda](www.rstudio.com)


#  TODO LIST

* how to initialize the COTS population??? [KTS developed a crude method using reef-level densities...]
* determine how exactly the coral/cots interaction will work? 
    + Can we get a maximal live coral cover for each grid cell? What determines this maximum
    + We have cover for two species of coral- how does this enter and influence the model?
    + Is the SDM just used to specify initial live coral cover? Or, does this SDM affect the
  model in a more long-lasting way?


# DATA DESCRIPTION
The population model uses data of three major types: 

1. GIS Rasters: 0.01 degrees (ca. 1 km) resolution. Approx 1.3 million pixels. We chose to use a 1 km resolution to represent geographic variation within the Great Barrier Reef.
We chose this resolution because it was the coarsest resolution that could successfully capture
the vast majority of the small reefs in the landscape.  
2. Population-level data: XYZ dataframe containing unique information (Z1, Z2, ...) about
each 1 km^2^ reef segment, or "population" (ca. 17k populations).  
3. Reef-level data: XYZ dataframe containing unique information (z1, Z2, ...) about 
each reef (variable sizes) (ca. 2k reefs)

## Read in master population-level data

```{r echo=FALSE}
###############################
# READ IN ENVIRONMENTAL COVARIATES
###############################

setwd(ENVDATA_DIRECTORY)
envData <- read.table("ENV_dataGBR.txt",header=T,sep="\t")
NPOPS <- nrow(envData)
    # generate a template grid
head(envData)

```

## Extract study region extent

```{r}
minx <- min(envData$x)-0.005  # farthest west
miny <- min(envData$y)-0.005  # farthest south
maxx <- max(envData$x)+0.005  # farthest east
maxy <- max(envData$y)+0.005  # farthest north
studyRegion <- extent(minx,maxx,miny,maxy)
cat(sprintf("Western margin is %s degrees, southern margin is %s degrees, eastern margin is %s degrees, and northern margin is %s degrees",minx,miny,maxx,maxy))

plot(studyRegion, main="Study Region (blank)")  

```

## Make template raster (to use as a template for further raster creation etc...)

```{r}
template <- raster(ext=studyRegion,resolution=0.01,vals=0)   # create empty raster
#plot(template)
setwd(SPATIALDATA_DIRECTORY)
writeRaster(template,filename="templateRaster.asc",format="ascii",overwrite=T)   # write to file
totalCells <- length(template@data@values)    # 1,335,372 cells

cat(sprintf("total cells in study region raster is %s",totalCells))
```

## Make reef ID raster
This GIS layer will be used in several downstream operations 

```{r}
   ## rasterize an environmental variable.
reefID <- as.numeric(envData$REEF_ID)
ndx <- !is.na(reefID)
reefIDraster <- rasterize(envData[,c('x','y')][ndx,], template, field=reefID[ndx])   # memory limitations

plot(reefIDraster)
```


## Visualize the reef polygon layer

```{r}
reefdatadir <- paste(SPATIALDATA_DIRECTORY,"\\Masked reefs",sep="")   
setwd(reefdatadir)
GBR_polygons <- readOGR(reefdatadir,layer="GBR_reefs")

plot(GBR_polygons)
```

## Compute the percent reef for each 1km raster cell
Camille generated a raster at 100m resolution from the GBR reef polygon layer. Here we use this new layer to compute the percent of each 1km raster cell that comprises actual reef. 

```{r}
reefraster <- ReadRaster("reefraster2.asc",projection=projection,plot=F)
#plot(reefraster)
reefraster <- reclassify(reefraster,rcl=c(NA,NA,0, -Inf,0.5,0, 0.6,Inf,1))  # temporary layer: reclassify to binary
compute_percentReef <- function(t,na.rm=TRUE) sum(t,na.rm)  ## aggregation function
reefpercent <- aggregate(reefraster, fact=10, fun=compute_percentReef, expand=TRUE, na.rm=TRUE) 
reefpercent <- reefpercent-1
plot(reefpercent)
setwd(SPATIALDATA_DIRECTORY)
writeRaster(reefpercent,filename="reefPercentRaster.asc",format="ascii",overwrite=T)   # write to file

#### determine reef percent for each cell of interest in the 1km grid
  # <- crop(x=reefpercent, y=template)
#reefpercent <- raster::resample(x=reefpercent, y=template, method="bilinear")        # first resample to make sure the grids line up properly (NOTE: need to specify from resample package, not popbio package!)
?extract
envData[,'PercentReef'] <- extract(reefpercent,envData[,c('x','y')])   # add a new column to the envData (XYZ format)


```

### Initial coral cover (GIS, population-level)
This layer is derived from several sources: XXX, XXX, XXX 




# RMARKDOWN EXAMPLES- FOR REFERENCE
inline equation: $A = \pi*r^{2}$

Here is an example of embedded R code:

```{r}
a<-2+2
a

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
